---
title: "Machine Learning"
author: "DJL"
date: "02/10/2016"
output:
  html_document:
    toc: yes
    toc_float: yes
---

#Overview

**Aim**  
Create Machine learning models using GAP data. 


#Libraries
```{r load_libs, tidy=TRUE}
rm(list = ls())
dev.off()
library(plyr)
library(dplyr)
library(randomForest)
library(lubridate)
library(doMC)
library(caret)
library(reshape)
library(tableone)
library(stargazer)
library(mice)
```

#Functions
```{r define functions, tidy=TRUE}

#heatmap plotting function
heatmap_ml<- function(model="NA",x_ax="NA",x_title="NA", y_ax="NA",y_title="NA",color_metric="NA",h_title="NA",save_output="NA") {
  model_data<-model$results
  model_data[,x_ax]<-as.factor(model_data[,x_ax])
  model_data[,y_ax]<-as.factor(model_data[,y_ax])
  model_data_metric<-model_data[,color_metric]
  print(str(model_data))
  ggplot(data = model_data, aes_string(x=x_ax, y=y_ax, fill=color_metric))+
    geom_tile(color = "white")+
    scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
      midpoint = mean(model_data_metric), limit = c(min(model_data_metric),max(model_data_metric)), space = "Lab", 
      name="") +
    theme_minimal()+ 
    theme(axis.text.x = element_text(angle = 90, vjust = 1, 
      size = 10, hjust = 1))+
    coord_fixed()+
    ggtitle(h_title) + xlab(x_title) + ylab(y_title)
  ggsave(save_output)
}

#find simple model. 
simple_model_fun<- function(ml_model,tol_percentage = 1, metric_sm=metric) {
    tolerance(ml_model$results, metric = metric_sm, tol = tol_percentage, maximize = TRUE)    
}

#helper Function for creation stratified by confusionmatrix result data frame when making tables of enviroment vs classification 
create_table_df<-function(input_data,use_columns,ml_data_predictions){
  #create object to be used from now on. 
  full_data<-input_data[,1:use_columns]
  #add rownames to predictions
  tpredictions<-data.frame(ml_data_predictions)
  rownames(tpredictions)<-full_data$sampleID


  #create test data with pheno predictions. 
  model_test_pheno<-merge(tpredictions,full_data,by.x="row.names",by.y="sampleID")
  ## Make confusion matrix column
  model_test_pheno<-model_test_pheno %>% mutate(ConfusionMatrix=ifelse(Phenotype == "FEP" & ml_data_predictions=="Control","FEP_misclassed",
                                            ifelse(Phenotype == "FEP" & ml_data_predictions=="FEP","FEP_true",
                                            ifelse(Phenotype != "FEP" & ml_data_predictions=="FEP","Control_misclassified","Control_true"))))
  # Center and Scale PRS
  model_test_pheno[20:28]<-scale(model_test_pheno[20:28])
  return (model_test_pheno)
}


# Function for making tables of confusion matrix stratified dataframes of ml models
table_fun_ml<- function(mldata,out_name_id,prs_col=(20:28)){
  #output name of model define
  t_title=out_name_id
  latex="_latex"
  latex="_latex"
  #Create Table of demographics with 4 predicted groups.
  listVars <- c("Gender","Age", "Ethnicity","BMI","Tobacco","ICD_DSM",names(mldata)[prs_col])
  catVars <- c("Ethnicity","Tobacco","ICD_DSM")
  table1 <- CreateTableOne(vars = listVars, data = mldata, factorVars = catVars,strata=c("ConfusionMatrix"),includeNA = T)
  table1print<-print(table1, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
  write.csv(table1print, file=paste(output_dir,project_id,"not_latex","_",t_title,"_","Table_CM_1_Demographics.csv",sep=""),row.names = TRUE, col.names = TRUE,quote=FALSE,sep=",")

  
  #Create Table of demographics with 4 predicted groups. 
  listVars2 <- c("Medication","dsmiv","icd10","ICD_DSM","PanssScore","PanssPositive","PanssNegative","PanssPsycho")
  catVars2 <- c("Medication","dsmiv","icd10","ICD_DSM")
  table2_full <- CreateTableOne(vars = listVars2, data = filter(mldata,Phenotype=="FEP"), factorVars = catVars2,strata=c("ConfusionMatrix"),includeNA = T)
  table2print<-print(table2_full,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
  write.csv(table2print, paste(output_dir,project_id,"not_latex","_",t_title,"_","Table_CM_1_meds_Demographics.csv",sep=""),row.names = TRUE, col.names = TRUE,quote=FALSE,sep = ",")
  
  #stargazer latex tables
  table1sg<-stargazer(table1print[,1:5],summary=FALSE, rownames=T)
  write.csv(table1sg,paste(output_dir,project_id,latex,"_",t_title,"_","Table_CM_1_Demographics.csv",sep=""),sep="",row.names=F,col.names = F)
  table2sg<-stargazer(table2print[,1:3],summary=FALSE, rownames=T)
  write.csv(table2sg,paste(output_dir,project_id,latex,"_",t_title,"_","Table_CM_1_meds_Demographics.csv",sep=""),sep="",row.names=F,col.names = F)
}


```


#Directories
```{r Define directories}
top_dir<-getwd()
data_dir <-"./P0_Characterise/output/"
data_dir0 <-"./data/"
output_dir <-"./P11_glmnet_bootstrap/output/"
figs_dir <-"./P11_glmnet_bootstrap/figs/"


#Pre and post fixes.
project_name = "glmnet_boot"
project_id = "p11_"

```




#Load Data
```{r Load data}

load(file=paste(data_dir,"GX_DF_adj_data.Rdata",sep=""))
Genesets_HD<-read.csv(paste(data_dir0,"GeneSetsMental_Pirooznia_2016.csv",sep=""))

```


#Imputation (PRS + Tobacco) + Demographic factors
```{r Load data}
##### Fix Polygenic Risk Score
#Set Seed
seed = 7
set.seed(seed)

#Factors of demographics
fa_ICD_DSM<-GX_DF_adj$ICD_DSM
fa_ICD_DSM[is.na(fa_ICD_DSM)]<- "Other_Psychosis"
GX_DF_adj$ICD_DSM<-fa_ICD_DSM
fa_Pheno <-as.factor(GX_DF_adj$Phenotype)
fa_Gender <-as.factor(GX_DF_adj$Gender)
fa_Ethnicity <-as.factor(GX_DF_adj$Ethnicity)
fa_Tobacco <-as.factor(GX_DF_adj$Tobacco)
#Model Matrix
design = model.matrix(~fa_Gender+fa_Ethnicity)  

#####Impute PRS
imputedesign = model.matrix(~fa_ICD_DSM+fa_Ethnicity)  #imputation factors

#create dataframe for input into impute
PRS_full_impute<-cbind(GX_DF_adj[,c("Phenotype","Ethnicity","ICD_DSM","PRS_0.1","sampleID")],imputedesign[,2:6])

#Impute with KNN using preprocess.
set.seed(seed)
ImputePCA<-preProcess(PRS_full_impute,method= c("knnImpute"))

#Get NAs
PRS_NAs<-PRS_full_impute[is.na(PRS_full_impute$PRS_0.1),]

#Predict 
resultspred<-cbind(predict(ImputePCA, PRS_full_impute)[4],PRS_full_impute[c(1:3,5)])

#Check all equal
all.equal(GX_DF_adj$sampleID[c(2:280,1)],resultspred$sampleID[c(2:280,1)])

#Prs back in original DF
GX_DF_adj$PRS_0.1<-resultspred$PRS_0.1

resultsNA<-resultspred[resultspred$sampleID%in%PRS_NAs$sampleID,]
resultspred$Imputed<-ifelse(resultspred$sampleID%in%PRS_NAs$sampleID,"Imputed", "Real")

#Make Plots
#imputed  vs PRS
title = "PRS_imputed_vs_real"
ggplot(data = resultspred, 
       aes(x = ICD_DSM, y=PRS_0.1  , colour = Imputed)) +
  geom_jitter() +
  geom_boxplot()+
  facet_wrap(~ Ethnicity)+
  ggtitle(title)
ggsave(paste(figs_dir,project_id,project_name,title,".png",sep=""))


```


####Define all 8 models with predictors 
```{r Create input dataframes for models}
#Define all 8 models with predictors 
Gx <-GX_DF_adj[,-c(1:3,5:27)]
Gx_Scz<-Gx[,c("Phenotype",names(GX_DF_adj)[names(GX_DF_adj)%in%Genesets_HD$Gene_names])]
PRS<-GX_DF_adj[,c(4,24)]  
Demographics<-cbind(GX_DF_adj[,c(4,6)],design[,2:5])
PRS_Demo<-cbind(GX_DF_adj[,c(4,6,24)],design[,2:5])
Gx_PRS <- cbind(PRS,GX_DF_adj[,-c(1:27)])
Gx_Demo <- cbind(Demographics,GX_DF_adj[,-c(1:27)])
Gx_Demo_PRS <- cbind(PRS_Demo,GX_DF_adj[,-c(1:27)])

#list of all model predictors
predictorlists<-list(Gx,Gx_Scz,PRS,Demographics,PRS_Demo,Gx_PRS,Gx_Demo,Gx_Demo_PRS)

#vector of model ids
modelID<-c("1_Gx","2_Gx_Scz","3_PRS","4_Demographics","5_PRS_Demo","6_Gx_PRS","7_Gx_Demo","8_Gx_Demo_PRS")


```



#Configure
```{r Define Cores and Seed}

#Allow 8 Cores
registerDoMC(cores = 8) 

set.seed(seed)
#Control for algorithms, Bootstrapping, 10k = ~3 hours 8 cores 5k predictors 280 samples.
control <- trainControl(method="boot", number=10000, savePredictions="all") ## Steve added sampling argument. READ!!


#Define Metric to choose best model
metric <- "Accuracy"

```


# Machine Learning
```{r Ml}
#GLMNET MODELS

##### Model 1
ModelNum=1
set.seed(seed)
glmnet_1_Gx <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_1_Gx, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)

##### Model 2
ModelNum=2
set.seed(seed)
glmnet_2_GxS <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_2_GxS, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)

##### Model 3
ModelNum=3
set.seed(seed)
glmnet_3_PRS <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glm", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_3_PRS, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)

##### Model 4
ModelNum=4
set.seed(seed)
glmnet_4_Dem <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_4_Dem, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)


##### Model 5
ModelNum= 5
set.seed(seed)
glmnet_5_PrsD <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_5_PrsD, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)

##### Model 6
ModelNum=6
set.seed(seed)
glmnet_6_Gx_PRS <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_6_Gx_PRS, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)

##### Model 7
ModelNum=7
set.seed(seed)
glmnet_7_Gx_dem <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_7_Gx_dem, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)

##### Model 8
ModelNum=8
set.seed(seed)
glmnet_8_Gx_d_prs <- train(Phenotype~., data=predictorlists[[ModelNum]], method="glmnet", metric=metric, preProc=c("center", "scale"), trControl=control) 
#Save Models
save(glmnet_8_Gx_d_prs, file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""), compress = T)



##### Load all saved models
for (ModelNum in 1:8){
  load(file=paste(output_dir,project_id,project_name,modelID[ModelNum],".rdata",sep=""))  
}


##### Make list of all and save
glmnet_models<-list(Gx = glmnet_1_Gx, Gx_Scz = glmnet_2_GxS, PRS = glmnet_3_PRS, Dem = glmnet_4_Dem, PRS_Dem = glmnet_5_PrsD, Gx_PRS = glmnet_6_Gx_PRS,  Gx_Dem = glmnet_7_Gx_dem,  Gx_Dem_PRS = glmnet_8_Gx_d_prs)
save(glmnet_models, file=paste(output_dir,project_id,project_name,"_all_models",".rdata",sep=""), compress = T)

```




## Check most important variables for models and plot tables
```{r}
load(file=paste(output_dir,project_id,project_name,"_all_models",".rdata",sep=""))



#Create loop to get Var imp and metrics
#Define List
glm_var_list<-NULL
#loop through models
for (ModelNum in 1:8){
  model_name<-modelID[ModelNum]
  glmvarimp<-varImp(  glmnet_models[[ModelNum]], scale = FALSE)
  glmvariables<-glmvarimp$importance
  glmvariables$Variables<-rownames(glmvariables)
  glmvariables$Model<-model_name
  glmvariables<-filter(glmvariables,Overall !=0)
  glmvariables<-arrange(glmvariables,desc(Overall))
  #colnames(glmvariables)<-c(model_name,"Variable")
  glm_var_list <- rbind(glm_var_list,glmvariables)
}
glm_var_list

write.csv(glm_var_list, file=paste(output_dir,project_id,project_name,"_glmnet_all_mod_predictors.csv",sep=""),row.names=F,quote=FALSE) #save variables

##Create loop to get metrics
#Resamples
ml_resample<-resamples(glmnet_models)

#Save Mean for final models
ml_resample_means<-sapply(ml_resample$values[,-1],mean)
Accuracy_index<-grep("*Accuracy",names(ml_resample_means))
Glmnet_metric_means<-as.data.frame(cbind(ml_resample_means[Accuracy_index],ml_resample_means[-Accuracy_index]))
#change col and rownames
colnames(Glmnet_metric_means)<-c("Accuracy","Kappa")
Glmnet_metric_means$Model<-modelID
#save
write.csv(Glmnet_metric_means[,c(3,1,2)], file=paste(output_dir,project_id,project_name,"_glmnet_all_mod_metrics.csv",sep=""),row.names=F,quote=FALSE)


#Define Metrics
use_metrics<-c("Accuracy","Kappa")

#For loop
for (pMetric in use_metrics){
  
  #Print Metric to make sure okay.
  print(pMetric)
  
  #Subset to Metric using grep
  Metric_grep = paste("*",pMetric,sep="") # define grep
  metric_results<-ml_resample$values[,grep(Metric_grep,colnames(ml_resample$values))] # grep subset
  names(metric_results)
  
  # Change model names. BE CAREFUL HERE!!!! THIS IS BASED ON ON A PREDEFINED LIST. IF YOU ARE UNSURE SWITCH TO THE FIRS LINE BELOW!!!!!
  #names(Accuracy_results)<-unlist(strsplit(x=names(Accuracy_results), split=".Accuracy",)) #use this if unsure. 
  names(metric_results)<-modelID # modelID !!!! BE CAREFUL !!!!

  #Sort models by metric
  metric_plot<-metric_results[,row.names(as.matrix(sort(sapply(metric_results,mean),decreasing = F)))]
  #add dataset name
  metric_plot$Datasets<-c("Glmnet_GAP")
  #Melt for ggplot
  metric_plot2<-melt(metric_plot, id="Datasets")  
  #Change colnames
  colnames(metric_plot2)<-c("GeneLists","Model",pMetric)

  #Boxplots
  title=paste(pMetric, "of 10k Bootstrap Glmnet Models in GAP",sep=" ")
  ggplot(metric_plot2,
    aes_string(x="Model",y=pMetric, colour = "Model")) +
    #geom_point() + 
    geom_boxplot(notch=T)+
    coord_flip()+
    ggtitle(title)
  ggsave(paste(figs_dir,project_id,project_name,"_",pMetric,"_boxplot.png",sep=""))
}



```

# Missclassification Functions
```{r missclassificationFunctions }
#Function 1: extract data
data_get_fun <-function(list_models,names_models){
  outputlist<- list() #define emptylist 
  #For loop start
  for (Mod_Num in 1:length(list_models)){
    tempMod<-list_models[[Mod_Num]]
    tempModname <- paste("DF",names_models[Mod_Num],sep="_")
    
    #Check if glmnet
    if (tempMod$method == "glmnet"){
      tempHP<-tempMod$bestTune
      tempDF<-tempMod$pred
      tempDF<-filter(tempDF, alpha == tempHP[,"alpha"]  & lambda == tempHP[,"lambda"])
      #tempDF$sampleID <- #add sampleID column to tempDF  
      
      #check else if GLM
    } else if (tempMod$method == "glm"){
      tempDF<-tempMod$pred
      #tempDF$sampleID <- #add sampleID column to tempDF  
    } #else #error if not glm or glmnet
      #print ("WARNING: input not recognised as caret glmnet or glm object")
      #tempDF<-NA
  
    #add to output list 
    outputlist[[tempModname]] <- tempDF #add tempDF to list
  
  } #end bracket for loop
  #return full output
  return(outputlist)
} #end bracket function 



## Table function 1 demographics
table_dem_fun<-function(pdata,table_name,stratify,listVars,catVars){
  table1 <- CreateTableOne(vars = listVars, data = pdata, factorVars = catVars,strata=c(stratify),includeNA = T)
  table1print<-print(table1)
  table1print<-table1print[,-length(names(data.frame(table1print)))]
  #write.csv(table1print, file=paste(P0_output_dir,"no_latex_",table_name,sep=""),row.names = TRUE, col.names = TRUE,quote=FALSE,sep = ",")
  #write.csv(stargazer(table1print,summary=FALSE), file=paste(P0_output_dir,"latex_",table_name,sep=""),row.names=F,col.names = F,sep="")
}


```


# Missclassification
```{r misclassification calculations}

##### Set up #####
#project name defined
project_name = "glmnet_missclass"
#Gene expression data to use. 
GX_data<-GX_DF_adj[1:27]

#Function to get dataframes
output_dfs<-data_get_fun(glmnet_models,modelID)


#Get Prediction boolean.
for (df_Num in 1:length(output_dfs)){  
  output_dfs[[df_Num]]<-mutate(output_dfs[[df_Num]], Prediction = pred == obs)  
}
save(output_dfs, file=paste(output_dir,project_id,project_name,"List_of_DFs_bool.rdata",sep=""), compress = T)

#Dataframe to add misclassification.
PhenoMC_Data<-GX_data
#Get percentages for each rowIndex and add to dataframe.
for (df_Num in 1:length(output_dfs)){  
  temp_percent<-output_dfs[[df_Num]] %>% group_by(rowIndex) %>% summarise(avg=mean(Prediction))
  #colnames(temp_percent)[2]<-names(output_dfs)[df_Num]
  colnames(temp_percent)[2]<-paste("Acc",modelID[df_Num],sep="_")
  PhenoMC_Data<-cbind(PhenoMC_Data,temp_percent[,2])
}

#Save output
write.csv(PhenoMC_Data, file=paste(output_dir,project_id,project_name,"_Sample_Percent_accuracy.csv",sep=""),row.names=F,quote=FALSE)
PhenoMC_Data<-read.csv(paste(output_dir,project_id,project_name,"_Sample_Percent_accuracy.csv",sep=""))

str(PhenoMC_Data)

```

#################### Output & Characterisation for best model
```{r output results for missclassification}

######################Density plots
#######All models
metric_plot2<-melt(PhenoMC_Data, id=c(names(PhenoMC_Data[,-c(28:35)])))
#Change colnames
colnames(metric_plot2)<-c(names(PhenoMC_Data[,-c(28:35)]),"Model","Percentage")
#Control vs FEP
title = "Density Plots Stratified by Control and FEP"
ggplot(metric_plot2, aes(Percentage, colour = Phenotype)) +
  geom_density(alpha=0.01)+
  facet_wrap(~ Model)+
  ggtitle(title)
ggsave(paste(figs_dir,project_id,project_name,"_Density_all_Pheno.png",sep=""))

#Control vs Schizophrenia vs Other Psychoses
title = "Density Plots Stratified by Control and Psychosis type"
ggplot(metric_plot2, aes(Percentage, colour = ICD_DSM)) +
  geom_density(alpha=0.01)+
  facet_wrap(~ Model)+
  ggtitle(title)
ggsave(paste(figs_dir,project_id,project_name,"_Density_all_diag.png",sep=""))

######Just Acc_1
names(PhenoMC_Data)
#Change colnames
#Control vs FEP
title = "Density Plot Stratified by Control and FEP"
ggplot(PhenoMC_Data, aes(Acc_1_Gx, colour = Phenotype)) +
  geom_density(alpha=0.01)+
  ggtitle(title)
ggsave(paste(figs_dir,project_id,project_name,"_1_Gx_Density_pheno.png",sep=""))

#Control vs Schizophrenia vs Other Psychoses
title = "Density Plots Stratified by Control and Psychosis type"
ggplot(PhenoMC_Data, aes(Acc_1_Gx, colour = ICD_DSM)) +
  geom_density(alpha=0.01)+
  ggtitle(title)
ggsave(paste(figs_dir,project_id,project_name,"_1_Gx_Density_diag.png",sep=""))


####### 1_Gx characterisation with phenotypes. 

##Categorical Variables
#1. Ethnicity
plot_fun_gx <-function(inputd,xax,yax,col,pfacets,ptitle) {
  ggplot(inputd, aes_string(x=xax,y=yax, colour = col)) +
    #geom_point() +  
    geom_jitter() +  
    geom_boxplot(notch=F)+
    coord_flip()+
    facet_wrap(as.formula(paste("~", pfacets)))+
    ggtitle(ptitle)  
}


plot_fun_gx(PhenoMC_Data,"ICD_DSM","Acc_1_Gx","ICD_DSM","Ethnicity","Titleofplot")


#2. Gender
plot_fun_gx(PhenoMC_Data,"ICD_DSM","Acc_1_Gx","ICD_DSM","Gender","Titleofplot")



#3. Smoking 
plot_fun_gx(PhenoMC_Data,"ICD_DSM","Acc_1_Gx","ICD_DSM","Tobacco","Titleofplot")

#4. Medication
plot_fun_gx(PhenoMC_Data,"ICD_DSM","Acc_1_Gx","ICD_DSM","Medication","Titleofplot")
plot_fun_gx(PhenoMC_Data,"Medication","Acc_1_Gx","Medication","ICD_DSM","Titleofplot")


## Continues Variables
#Create bins
?quantile
?cut
PhenoMC_Data$Cuts<-cut(PhenoMC_Data$Acc_1_Gx,quantile(PhenoMC_Data$Acc_1_Gx,probs = c(0, 20, 40, 60, 80,100)/100))
quantile(PhenoMC_Data$Acc_1_Gx,probs())
quantile(x<-c(1:100,1:100),  probs = c(25, 50, 1, 2, 5, 10, 50, NA)/100)
quantile(x<-c(0:100,0:100),  
quantile(PhenoMC_Data$Acc_1_Gx,probs = c(0, 20, 40, 60, 80,100)/100)



#5. Age 
plot_fun_gx(PhenoMC_Data,"ICD_DSM","Acc_1_Gx","ICD_DSM","Medication","Titleofplot")
plot_fun_gx(PhenoMC_Data,"Cuts","Age","Cuts","ICD_DSM","Titleofplot")
ggplot(PhenoMC_Data, aes(Acc_1_Gx, colour = ICD_DSM)) +
ggplot(PhenoMC_Data, aes(x=Cuts,y=Age, colour = ICD_DSM)) +
  geom_jitter() +  
  geom_boxplot(notch=F)+
  coord_flip()+
  ggtitle(title)



#6. PRS (not imputed.)
PhenoMC_Data_white<-filter(PhenoMC_Data,Ethnicity=="White")
PhenoMC_Data_white$Cuts<-cut(PhenoMC_Data_white$Acc_1_Gx,quantile(PhenoMC_Data_white$Acc_1_Gx,probs = c(0, 20, 40, 60, 80,100)/100))
plot_fun_gx(PhenoMC_Data_white,"ICD_DSM","PRS_0.1","ICD_DSM","Cuts","Titleofplot")
PhenoMC_Data_white$Cuts


```





## Under Construction ##########################################################################################
```{r}
###################################################################################
## Under Construction
#pick model and name
input_model=glmnet_fit_b
input_model_name="glm"
#use model to predict training data
predict_fit<-predict.train(input_model, newdata = GX_DF_full_adj)
cmatrix<-confusionMatrix(predict_fit, GX_DF_full_adj$Phenotype)
matrixout<-data.frame(c(cmatrix$overall,cmatrix$byClass))
#save metrics
write.table(t(matrixout), file=paste(output_dir,project_id,project_name,"glmnet_confusionout.csv",sep=""),row.names=F,quote=FALSE,sep = ",")


#make dataframes
model_strat_confuse_df<-create_table_df(GX_DF_full_adj,27,predict_fit)
#plot tables
table_fun_ml(model_strat_confuse_df,input_model_name)


#modellist<- list(HF_glm=p1_glm,HD_glm=p2_glm)

#p1data<-p1_glm$pred
#str(p1_glm$pred)
#table(p1data$lambda)

#p1filter<-filter(p1data,alpha== 0.1)
#p1filter<-filter(p1filter,lambda < 0.0105)
dim(p1filter)
table(p1filter$rowIndex)

smallp1<-p1filter[!duplicated(p1filter$rowIndex),]
dim(smallp1)
smallsort<-arrange(smallp1,rowIndex)

cbind(smallsort[,1:3],expressionRFE$Phenotype)


```
